### 2024-04-28 thesis/chapters/04-implementation.tex
Turn outline into section headings

\chapter{Implementation} \label{section:implementation} 
The methodology applied so far for the dataset recorded in a laboratory is to be tested on vibration signals from the industrial environment. Established guidelines for introducing a monitoring system have been described in the section on technical standards (Section~\ref{section:technical-standards}).

We employ a slightly customized validation procedure that involves the selection of machines to be monitored, identifying positions for measurements according to technical specifications, taking preliminary readings, and developing a device capable of capturing the usual failure modes. The collection of the novel dataset is accompanied by an agreed upon schedule.

\section{Machinery for monitoring}
Three kinds of rotating machines are at disposal for vibration measurements. Those are a standing fan, scroll compressor, and water pump.

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.17\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/machine-fan.jpg}
        \caption{\footnotesize Fan}
        \label{fig:machine:fan}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.17\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/machine-compressor.jpg}
        \caption{\footnotesize Compressor}
        \label{fig:machine:compressor}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.31\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/machine-ksb-pump.jpg}
        \caption{\footnotesize Water pump KSB}
        \label{fig:machine:pump-ksb}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.31\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/machine-sigma-pump.jpg}
        \caption{\footnotesize Water pump Sigma}
        \label{fig:machine:pump-sigma}
    \end{subfigure}
    \caption{Machines dedicated for vibration measurements}
\end{figure}

\textbf{Standing fan} is model \emph{Kalorik~TKG~VT1037} (Fig.~\ref{fig:machine:fan}) and one unit is available to us. It serves as a test bed during the sensor unit development. The sensor is placed on the plastic casing at the back of the drive motor. The fan has a 45~cm diameter with 3~propellers and a power of 45~W (class~\rom{1}). It has a switch for 3 rotational speeds, which are approximately 18.5~Hz (1100~rpm), 20.4~Hz (1200~rpm), and 22~Hz (1300~rpm). Acceleration observed is $\pm 5\;\mathrm{m/s}^2$ in radial and tangential directions, and $\pm 1\;\mathrm{m/s}^2$ in axial direction.

\textbf{Scroll compressor} is model \emph{Copeland~ZR16} (Fig.~\ref{fig:machine:compressor}). Altogether, two units are available in two independent air conditioning units for the data center. The compressor has 9.7~kW of power (class~\rom{1}) and rotates at 2900 rpm (48.3 Hz). Two possible measurement locations are located on the sides on top of the bearings, just above the base and above the scroll. Accelerations picked up by the sensor are $\pm 3\;\mathrm{m/s}^2$ tops.

\textbf{Water pumps} are available as 3 units in municipal drinking water pumping station. The apparatus consists of a single-stage axially split volute casing pump and an attached electric induction motor.

The newer primary pumps with bundled wireless cloud monitoring system are two units of \emph{KSB Omega~300-560} (Fig.~\ref{fig:machine:pump-ksb}). The pumps were commissioned in 2018, and they rotate at 1493 rpm (24.9 Hz). The electric motor provides 400~kW of power (class~\rom{3}). The outer bearing on the pump vibrates with an acceleration of $\pm 5 \;\mathrm{m/s}^2$.

The secondary pump is one unit named \emph{Sigma~300-OVD-600} (Fig.~\ref{fig:machine:pump-sigma}) installed in 1986. It rotates at 1485 rpm (24.75 Hz), and its electric motor has power of 450~kW (class~\rom{3}). Each pump and motor have 2 bearings. Therefore, there are 4 measurement positions in total.

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/ksb-guard-rms.png}
        \caption{Vibration rms velocity dashboard}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/ksb-guard-spectrum.png}
        \caption{Vibration spectral analysis dashboard}
    \end{subfigure}
     \caption{KSB Guard cloud monitoring for pumps}
     \label{fig:design:ksb-guard}
\end{figure}

\textbf{Complete schedule} for planned vibration data gathering on the designated machines and measurement procedure is described in Appendix~\ref{appendix:technical-docs}. In addition, we are able to export the entire historical record at hourly intervals. Logs from \textbf{KSB Guard} cloud monitoring tool contain vibration rms velocities and frequency spectra for two monitored KSB water pumps (Figure~\ref{fig:design:ksb-guard}).

\section{Sensor hardware and drivers}
In preliminary exploratory vibration measurements of machines, the ADXL335 accelerometer is connected to BeagleBone Black microcontroller. The bandwidth, sensitivity, and sampling frequency of the makeshift solution do not appear to be adequate (Table~\ref{tab:design:hw-sensors}).

\begin{table}[ht]
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{|l|l|l|}
\hline
\textbf{Accelerometer}                           & \textbf{ADXL335} & \textbf{IIS3DWB}   \\ \hline
\textbf{Vendor}                                  & Analog Devices   & STMicroelectronics \\ \hline
\textbf{Bus}                                     & Analog           & SPI                \\ \hline
\textbf{Axis}                                    & 3                & 1 or 3             \\ \hline
\textbf{Range} (g)                               & $\pm$ 3          & $\pm$ 2 to 16      \\ \hline
\textbf{Bandwidth} (kHz)                         & 0.55 - 1.6       & 5 - 6.3            \\ \hline
\textbf{Sensitivity} (mg/LSB)                    & 2.930            & 0.061       \\ \hline
\textbf{Noise density} ($\mu g / \sqrt{\mathrm{Hz}}$ rms) & 150 - 300        & 75                 \\ \hline
\textbf{Microcontroller}                         & Beaglebone Black & ESP32-PoE-ISO      \\ \hline
\textbf{CPU SoC}                                 & TI Sitara AM3358 & ESP32-WROOM-32     \\ \hline
\textbf{Output data rate} (kHz)                  & 2.5              & 26.7               \\ \hline
\textbf{ADC resolution} (bit)                    & 12               & 16                 \\ \hline
\textbf{FIFO}                                    & -                & 3 kB (512 samples) \\ \hline
\end{tabular}
\caption{Hardware parameters of previous and proposed sensor units}
\label{tab:design:hw-sensors}
\end{table}

The IIS3DWB accelerometer is a cheap enough MEMS accelerometer that approaches industrial standards for vibration monitoring. ESP32-PoE-ISO is chosen as a microcontroller development kit because it has an SD card slot connected via SD/MMC bus. It enables us to store all samples from FIFO. The accelerometer uses an SPI bus with a maximum speed of 10 MHz that can be connected to any physical GPIO pin. The block diagram of the designed hardware device is in Figure~\ref{fig:design:block-diagram-hw}.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{assets/design/hw-block-schematic.png}
	\caption{Sensor unit hardware block diagram}
	\label{fig:design:block-diagram-hw}
\end{figure}

The drivers for necessary low-level interfaces and FAT32 filesystem are  implemented already in ESP-IDF SDK\footnote{ESP-IDF SDK: \url{https://docs.espressif.com/projects/esp-idf/en/latest/esp32/api-reference/index.html}}. The accelerometer driver is also made available by the vendor\footnote{IIS3DWB driver: \url{https://github.com/STMicroelectronics/iis3dwb-pid}}. Activity diagrams of required firmware functionality are in Appendix~\ref{appendix:technical-docs}.

\section{Preliminary measurements}
During the initial on-site inspections of machinery, we carried out preliminary measurements of vibration signals. Raw waveforms of acceleration for the fan, two different compressors, and the outer bearing on a pump and motor in three axes are plotted in Figure~\ref{fig:design:preliminary-acceleration}. The acceleration rarely exceeds 1~g.

In frequency spectra in Figure~\ref{fig:design:preliminary-spectrum} behavior of machines can be clearly discerned below the resonance frequency of the accelerometer (550 Hz). The compressor suspected to be faulty has a stronger \nth{4} harmonic and above than the other one.

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/design/EDA-custom-dataset-temporal.png}
        \caption{Acceleration in $\mathrm{m/s}^2$}
        \label{fig:design:preliminary-acceleration}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/design/EDA-custom-dataset-velocity.png}
        \caption{Approximate velocity in $\mathrm{mm/s}$}
        \label{fig:design:preliminary-velocity}
    \end{subfigure} 
    \caption{Temporal domain waveforms in all spatial directions with 300 ms duration at 1 s timestamp}
\end{figure}

Figure~\ref{fig:design:preliminary-velocity} documents the effort to obtain vibration velocity (as per ISO~20186) by integration with the composite trapezoidal rule. The significant drift after integration is eliminated by subtracting the mean of the envelope constructed using cubic interpolation of peaks. Envelopes must be tweaked further to achieve more accurate values. The orbital plot of the first harmonic frequency $f_{max} \pm 5\;\mathrm{Hz}$ in Figure~\ref{fig:design:orbital-analysis} allows us to check the sensor deviation out of the axis of rotation.

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{assets/design/compressor-orbitals-1x.png}
		\caption{Orbital analysis of compressor's \nth{1}~harmonic frequency}
		\label{fig:design:orbital-analysis}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.54\textwidth}
        \includegraphics[width=\textwidth]{assets/design/EDA-custom-dataset-spectral-X-axis.png}
        \caption{FFT with Hann window of length $2^{10}$ ($\approx$~409~ms)}
        \label{fig:design:preliminary-spectrum}
    \end{subfigure} 
    \caption{Orbital and spectral methods of vibration signal analysis}
\end{figure}

>>>>

\chapter{Implementation} \label{section:implementation}

\section{Data analysis}
% Jupyter notebook - TSFEL, numpy, pandas (data handling),  scipy (signal processing and stats),   | sklearn, imblearn | matplotlib

\section{Firmware}
% In C, ESP-IDF SDK and FreeRTOS tasks
% Found driver online
% Problems with timing - binary write (osciloscope pictures)
% tool for convert bin2csv

### 2024-04-28 thesis/chapters/06-conclusion.tex
Turn outline into section headings

\chapter{Conclusion} \label{section:conclusion}  
In the thesis we focused on trend indicator selection for an inexpensive industrial condition monitoring solution from vibration signals. The goal is to enable timely fault detection of machinery parts with as little input data as possible. For that purpose, we answer four research questions.

Attributes extracted to describe machine behavior come mainly from descriptive statistics, audio signal processing studies, and vibrodiagnostics technical standards (ISO 20186 and ISO 13343). These formulas compute 10 features summarizing the waveform in the temporal domain and 11 features characterizing spectral density estimation in 3 spatial directions \textbf{(RQ1)}.

In order to achieve more pronounced data savings, we choose the subset of 3 features in each domain by keeping the ones with the most similarity to the target variable. Feature selection metrics of the correlation coefficient, F statistic, mutual information, and their rank product are applicable in supervised learning. The features are squished from multiple dimensions using the Euclidian norm. Lossy compression ratios attained are 2381:1 for all features and 25000:1 for 6 features in the MaFaulDa dataset. We managed to discard more than 99.995\% of irrelevant data \textbf{(RQ2)}.

Feature subsets are subjected after normalization to a k-nearest neighbor classifier that ascertains their relative fault detection power. Because of model overtraining with small k, the feature triplets equal or slightly outperformed the whole set of features on the validation set with accuracy up to 10\%. The spectral features reach higher accuracies than temporal because of their smaller interdependency \textbf{(RQ3)}.

The ensemble of feature selection with rank product produces the best model performance out of three combined in the majority of situations. No approach could find a triplet of predictors with an accuracy close to optimal one discovered exhaustively. Training on three principal components produced better accuracy than filtering feature selection, but PC mapping onto original trend indicators is unclear even in loading plots \textbf{(RQ3)}.

The considerable obstacle in an autonomous fault detection system deployment is the availability of labels for target variables. Annotations can be assigned belatedly or even never. Incremental learning kNN model on an unbalanced dataset on the whole feature set achieves at best 90\% accuracy with immediate feedback, 85\% with labels coming in 250 long tumbling windows, and 82\% with just 25\% of observations associated with the label. The comparable model trained in batch reaches an accuracy of 98\% \textbf{(RQ4)}.

Conclusions so far have been made on the MaFaulDa dataset imitating the realistic conditions. Therefore, in proper validation of the proposed solution in practice, we will compare it with the custom-made dataset.

The vibration signals will be gathered on compressors in air conditioning units and water pumps in municipal pumping stations. We will develop firmware for sensor units capable of sampling accelerometers and saving measurements onto SD cards. The challenge awaits us in labeling samples themselves as it requires substantial expert knowledge.

Feature selection methods should be incorporated into incremental learning. Hyperparameters shall be adjusted, and the balancing method used to increase model performance. The outlined tasks awaiting completion are planned for the DP~\rom{3} phase.

>>>>

\chapter{Conclusion} \label{section:conclusion}
